
---
title: "Part-II Complex Model"
author: "Chenxi Wu"
geometry: margin=1in
date: "11/30/2019"
output: 
    pdf_document:
        highlight: pygment
        toc: false
        toc_depth: 2
        fig_caption: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, 
                      warning=FALSE, 
                      message=FALSE,
                      fig.align="center",
                      fig.pos='H', 
                      results="asis")
```

```{r packages, echo = F, message = F, warning = F}
library(tidyverse)
library(knitr)
library(gbm)
library(xtable)
library(kableExtra)
library(broom)
library(car)
library(GGally)
library(caret)
library(randomForest)
library(BAS)
```


```{r read-data, echo = F, message = F, warning = F}
load("paintings_train.Rdata")
load("paintings_test.Rdata")
```



```{r data cleaning, echo = F, message = F, warning = F}
# merge the two dataset first
paint <-rbind(paintings_train, paintings_test)
# Fix position > 1
paint$position[paint$position > 1] <- paint$position[paint$position > 1]/100
# regroup some categorical variables
paint <- paint %>%
  mutate(winningbiddertype = as.factor(dplyr::recode(na_if(winningbiddertype,""),
                                              .missing = 'U')),
         dealer = as.factor(dplyr::recode(dealer, 
                                          'L' = 'LP', 
                                          'P' = 'LP')), 
         endbuyer = as.factor(dplyr::recode(na_if(endbuyer,""),
                                     .missing = 'EU',
                                     'E' = 'EU', 
                                     'U' = 'EU',
                                     'B' = 'C')),
         Interm = as.factor(dplyr::recode(Interm,
                                   .missing = 0,
                                   `0` = 0,
                                   `1` = 1)),
         authorstyle = dplyr::recode(authorstyle,
                                      "n/a" = 0,
                                      .default = 1),
         type_intermed = as.factor(dplyr::recode(na_if(type_intermed,''),
                                          .missing='Unknown')),
         Shape = as.factor(dplyr::recode(na_if(Shape,""),
                                  .missing = 'other',
                                  'squ_rect' = 'squ_rect',
                                  .default = 'other')),
         materialCat = as.factor(dplyr::recode(na_if(materialCat,""),
                                        .missing = 'other')),
         school_pntg = as.factor(dplyr::recode(school_pntg,
                                               'S' = 'X',
                                               'A' = 'X',
                                               'G' = 'X')), 
         origin_cat = as.factor(dplyr::recode(origin_cat,
                                    "S" = "O")), 
         nfigures = as.factor(dplyr::recode(nfigures,
                                            `0` = 0,
                                            .default = 1)),
         price = as.numeric(str_remove_all(paint$price, ","))) %>% 
  # change variables into appropriate format
  mutate_at(vars(dealer, origin_author, origin_cat, school_pntg, 
                 diff_origin, artistliving, authorstyle, 
                 winningbiddertype, endbuyer, Interm, type_intermed, 
                 material, mat, materialCat, Shape, engraved, original, 
                 prevcoll, othartist, paired, figures, finished, lrgfont, 
                 relig, landsALL, lands_sc, lands_elem, lands_figs, lands_ment, 
                 arch, mytho, peasant, othgenre, singlefig, portrait, 
                 still_life, discauth, history, allegory, pastorale,other), as.factor) 
# impute Surface on different groups
paint <- paint %>% mutate(Key = seq(1:nrow(paint)))
paint1 <- paint %>% filter(materialCat == 'canvas', relig == 1) %>% 
  mutate(Surface = dplyr::recode(Surface,
                          `0` = median(Surface, na.rm = T),
                          .missing = median(Surface, na.rm = T)))
paint2 <- paint %>% filter(materialCat == 'canvas', relig == 0) %>%
  mutate(Surface = dplyr::recode(Surface,
                                 `0` = median(Surface, na.rm = T),
                          .missing = median(Surface, na.rm=T)))
paint3 <- paint %>% filter(materialCat == 'copper', relig == 1) %>% 
  mutate(Surface = dplyr::recode(Surface,
                                 `0` = median(Surface, na.rm = T),
                          .missing = median(Surface, na.rm = T)))
paint4 <- paint %>% filter(materialCat == 'copper', relig == 0) %>%
  mutate(Surface = dplyr::recode(Surface,
                                 `0` = median(Surface, na.rm = T),
                          .missing = median(Surface, na.rm=T)))
paint5 <- paint %>% filter(materialCat == 'other', relig == 1) %>% 
  mutate(Surface = dplyr::recode(Surface,
                                 `0` = median(Surface, na.rm = T),
                          .missing = median(Surface, na.rm = T)))
paint6 <- paint %>% filter(materialCat == 'other', relig == 0) %>%
  mutate(Surface = dplyr::recode(Surface,
                                 `0` = median(Surface, na.rm = T),
                          .missing = median(Surface, na.rm=T)))
paint7 <- paint %>% filter(materialCat == 'wood', relig == 1) %>% 
  mutate(Surface = dplyr::recode(Surface,
                                 `0` = median(Surface, na.rm = T),
                          .missing = median(Surface, na.rm = T)))
paint8 <- paint %>% filter(materialCat == 'wood', relig == 0) %>%
  mutate(Surface = dplyr::recode(Surface,
                                 `0` = median(Surface, na.rm = T),
                          .missing = median(Surface, na.rm=T)))
paint <- rbind(paint1, paint2, paint3, paint4, paint5, paint6, paint7, paint8) %>% 
  arrange(Key) %>% 
  select(-Key) %>% 
  dplyr::select(-c("sale", "lot", "count","subject", "author", "origin_author", 
                   "winningbidder","type_intermed","material","mat", "still_life", "portrait",
                   "Diam_in","Surface_Rnd","Surface_Rect","Width_in","Height_in","other",
                   "pastorale","history","lands_ment","original","allegory","position",
                   "school_pntg", "winningbiddertype", "figures", "singlefig"))
paint$Surface <- log(paint$Surface)
# paint$year_sq <- (paint$year)^2
# paint$year_cu <- (paint$year)^3
# summary(paint)
remove <- c("attributed to ", "after ", "in the manner of ", "in the taste of ", "in the genre of ", "in the style of ")
paint$authorstandard <- str_remove_all(paint$authorstandard, 
                                             paste(remove, collapse = "|"))
# Split back into train and test
paint_train <- paint %>% filter(!is.na(logprice))
paint_test <- paint %>% filter(is.na(logprice))
level <- c(1, 0)
train_author <- paint_train %>% 
  group_by(authorstandard) %>% 
  summarise(avg_price = mean(price)) %>% 
  arrange(desc(avg_price)) %>% 
  mutate(expensive = case_when(avg_price >= 2000 ~ level[1],
                           avg_price < 2000 ~ level[2]))
ggplot(data = train_author, aes(y = avg_price, x = 1:473, col = as.factor(expensive))) +
  geom_point() +
  labs(title = "Average price for each author", x = "Rank", y = "Average price")
# attribute each author name to its group
author_group = map(level, ~train_author$authorstandard[train_author$expensive == .x]) %>%
  set_names(level)
# group the training set
paint_train <- paint_train %>% 
  mutate(expensive = case_when(authorstandard %in% author_group[[1]] ~ level[1], 
                           authorstandard %in% author_group[[2]] ~ level[2]) %>% 
           as.factor())
# group the test set using same metric, set new authors to 0
paint_test <- paint_test %>% 
  mutate(expensive = case_when(authorstandard %in% author_group[[1]] ~ level[1], 
                           authorstandard %in% author_group[[2]] ~ level[2]) %>% 
           as.factor)
paint_test$expensive[is.na(paint_test$expensive)] <- level[2]
```



```{r final model, fig.asp=1}
new <- lm(logprice ~ year + dealer + expensive + authorstyle + endbuyer + Interm + Surface + finished + lrgfont + diff_origin, data = paint_train)
final_model <- new
```

### Predictions of Validation set and Top 10 paintings
```{r}
#make same changes to validation set
load("paintings_validation.RData")
# Fix position > 1
paintings_validation$position[paintings_validation$position > 1] <- paintings_validation$position[paintings_validation$position > 1]/100
# regroup some categorical variables
paintings_validation <- paintings_validation %>%
  mutate(winningbiddertype = as.factor(dplyr::recode(na_if(winningbiddertype,""),
                                              .missing = 'U')),
         dealer = as.factor(dplyr::recode(dealer, 
                                          'L' = 'LP', 
                                          'P' = 'LP')), 
         endbuyer = as.factor(dplyr::recode(na_if(endbuyer,""),
                                     .missing = 'EU',
                                     'E' = 'EU', 
                                     'U' = 'EU',
                                     'B' = 'C')),
         Interm = as.factor(dplyr::recode(Interm,
                                   .missing = 0,
                                   `0` = 0,
                                   `1` = 1)),
         authorstyle = dplyr::recode(authorstyle,
                                      "n/a" = 0,
                                      .default = 1),
         type_intermed = as.factor(dplyr::recode(na_if(type_intermed,''),
                                          .missing='Unknown')),
         Shape = as.factor(dplyr::recode(na_if(Shape,""),
                                  .missing = 'other',
                                  'squ_rect' = 'squ_rect',
                                  .default = 'other')),
         materialCat = as.factor(dplyr::recode(na_if(materialCat,""),
                                        .missing = 'other')),
         school_pntg = as.factor(dplyr::recode(school_pntg,
                                               'S' = 'X',
                                               'A' = 'X',
                                               'G' = 'X')), 
         origin_cat = as.factor(dplyr::recode(origin_cat,
                                    "S" = "O")), 
         nfigures = as.factor(dplyr::recode(nfigures,
                                            `0` = 0,
                                            .default = 1))) %>%
  # change variables into appropriate format
  mutate_at(vars(dealer, origin_author, origin_cat, school_pntg, 
                 diff_origin, artistliving, authorstyle, 
                 winningbiddertype, endbuyer, Interm, type_intermed, 
                 material, mat, materialCat, Shape, engraved, original, 
                 prevcoll, othartist, paired, figures, finished, lrgfont, 
                 relig, landsALL, lands_sc, lands_elem, lands_figs, lands_ment, 
                 arch, mytho, peasant, othgenre, singlefig, portrait, 
                 still_life, discauth, history, allegory, pastorale,other), as.factor) 
# impute Surface on different groups
paintings_validation <- paintings_validation %>% mutate(Key = seq(1:nrow(paintings_validation)))
paint1 <- paintings_validation %>% filter(materialCat == 'canvas', relig == 1) %>% 
  mutate(Surface = dplyr::recode(Surface,
                          `0` = median(Surface, na.rm = T),
                          .missing = median(Surface, na.rm = T)))
paint2 <- paintings_validation %>% filter(materialCat == 'canvas', relig == 0) %>%
  mutate(Surface = dplyr::recode(Surface,
                                 `0` = median(Surface, na.rm = T),
                          .missing = median(Surface, na.rm=T)))
paint3 <- paintings_validation %>% filter(materialCat == 'copper', relig == 1) %>% 
  mutate(Surface = dplyr::recode(Surface,
                                 `0` = median(Surface, na.rm = T),
                          .missing = median(Surface, na.rm = T)))
paint4 <- paintings_validation %>% filter(materialCat == 'copper', relig == 0) %>%
  mutate(Surface = dplyr::recode(Surface,
                                 `0` = median(Surface, na.rm = T),
                          .missing = median(Surface, na.rm=T)))
paint5 <- paintings_validation %>% filter(materialCat == 'other', relig == 1) %>% 
  mutate(Surface = dplyr::recode(Surface,
                                 `0` = median(Surface, na.rm = T),
                          .missing = median(Surface, na.rm = T)))
paint6 <- paintings_validation %>% filter(materialCat == 'other', relig == 0) %>%
  mutate(Surface = dplyr::recode(Surface,
                                 `0` = median(Surface, na.rm = T),
                          .missing = median(Surface, na.rm=T)))
paint7 <- paintings_validation %>% filter(materialCat == 'wood', relig == 1) %>% 
  mutate(Surface = dplyr::recode(Surface,
                                 `0` = median(Surface, na.rm = T),
                          .missing = median(Surface, na.rm = T)))
paint8 <- paintings_validation %>% filter(materialCat == 'wood', relig == 0) %>%
  mutate(Surface = dplyr::recode(Surface,
                                 `0` = median(Surface, na.rm = T),
                          .missing = median(Surface, na.rm=T)))
paintings_validation <- rbind(paint1, paint2, paint3, paint4, paint5, paint6, paint7, paint8) %>% 
  arrange(Key) %>% 
  select(-Key) %>% 
  dplyr::select(-c("sale", "lot", "count","subject", "author", "origin_author",
                   "winningbidder","type_intermed","material","mat", "still_life", "portrait",
                   "Diam_in","Surface_Rnd","Surface_Rect","Width_in","Height_in","other",
                   "pastorale","history","lands_ment","original","allegory","position",
                   "school_pntg", "winningbiddertype", "figures", "singlefig"))
paintings_validation$Surface <- log(paintings_validation$Surface)
remove <- c("attributed to ", "after ", "in the manner of ", "in the taste of ", "in the genre of ", "in the style of ")
paintings_validation$authorstandard <- str_remove_all(paintings_validation$authorstandard, 
                                             paste(remove, collapse = "|"))
level <- c(1, 0)
train_author <- paint_train %>% 
  group_by(authorstandard) %>% 
  summarise(avg_price = mean(price)) %>% 
  arrange(desc(avg_price)) %>% 
  mutate(expensive = case_when(avg_price >= 2000 ~ level[1],
                           avg_price < 2000 ~ level[2]))
# attribute each author name to its group
author_group = map(level, ~train_author$authorstandard[train_author$expensive == .x]) %>%
  set_names(level)
# group the training set
paint_train <- paint_train %>% 
  mutate(expensive = case_when(authorstandard %in% author_group[[1]] ~ level[1], 
                           authorstandard %in% author_group[[2]] ~ level[2]) %>% 
           as.factor())
# group the test set using same metric, set new authors to 0
paint_test <- paint_test %>% 
  mutate(expensive = case_when(authorstandard %in% author_group[[1]] ~ level[1], 
                           authorstandard %in% author_group[[2]] ~ level[2]) %>% 
           as.factor)
paint_test$expensive[is.na(paint_test$expensive)] <- level[2]
paintings_validation <- paintings_validation %>% 
  mutate(expensive = case_when(authorstandard %in% author_group[[1]] ~ level[1],
                              authorstandard %in% author_group[[2]] ~ level[2]) %>% as.factor)
paintings_validation$expensive[is.na(paintings_validation$expensive)] <- level[2]
```

```{r}
#predict top 10
fitted <- exp(predict(final_model, newdata = paintings_validation))
Top10 <- cbind(paintings_validation, fitted) %>% arrange(desc(fitted)) %>% slice(1:10) %>% select(-c(logprice,price,authorstandard))
kable(Top10, digits = 3)
```
Using our model for predicting price for validation data set, we got our top 10 valuable paintings. We can see that these paintings all share certain important features, such as they are all from the same dealer, R. In addition, endbuyers are mostly from category C, the dealer devotes an additional paragraph and an intermediary is involved in the transaction etc. This is quite expected due to the way we constructed our model. 

```{r}
predictions_validation = as.data.frame(
  exp(predict(final_model, newdata=paintings_validation, 
              interval = "pred")))
save(predictions_validation, file="prediction-validation.Rdata")
```



